---
title: "03_eating_disorder_stats"
author: "SP"
date: "`r Sys.Date()`"
output: 
  workflowr::wflow_html: 
    df_print: paged
    fig_width: 11
    fig_height: 8
    fig_caption: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = FALSE)
# knitr::opts_knit$set(root.dir = dirname(getwd()))
```

```{r libs, include=FALSE}
do_source_data_creation <- TRUE
source("code/eating_disorder_data_proc.R")
load_inst_pkgs("tidyverse", "psych", "janitor", "knitr")
```

# Sample periods
- Pre-Covid from `r study_start` until `r covid_start`: `r pre_cov_dur` years
- During Covid pandemic until `r study_end`: `r cov_dur` years


# sample characteristics
## number of diagnosis per case
```{r}
df_diag %>%
  summarise(n = n_distinct(icd_code), .by = case_id) %>%
  ggplot(aes(x = n)) + geom_bar() + ggtitle("number of diagnoses per case")
```

## number of cases per year pre/during covid
```{r}
df_ed %>% 
  group_by(covid_pan) %>%
  summarise(n_cases = n_distinct(case_id),
            n_patients = n_distinct(p_id)) %>%
  mutate(years = c(pre_cov_dur, cov_dur)) %>%
  adorn_totals() %>%
  mutate(n_cases_year = n_cases/years,
         n_patients_year = n_patients/years)
```

## mean & sd
```{r}
summarise_mean_sd <- function(df) {
  summarise(df, 
            # across(where(~is.numeric(.x) | is.POSIXct(.x)), list(mean = ~mean(., na.rm = FALSE), sd = SD)), 
            across(c(age_adm, length_stay), list(mean = ~mean(., na.rm = TRUE), sd = SD)), 
            across(sex, list(mean = ~mean(as.numeric(.)-1, na.rm = TRUE), sd = ~SD(as.numeric(.)-1))),
            n_cases = n())
}

df_ed_stats <- 
  # F50 all
  df_ed %>% summarise_mean_sd() %>% mutate(covid_pan = "total_dur", .before = 1) %>% 
  bind_rows(df_ed %>% group_by(covid_pan) %>% summarise_mean_sd()) %>% 
  mutate(icd_f50 = "all_diag", .before = 2) %>% 
  # F50 groups
  bind_rows(df_ed %>% group_by(icd_f50) %>% summarise_mean_sd() %>% mutate(covid_pan = "total_dur", .before = 1)) %>% 
  bind_rows(df_ed %>% group_by(icd_f50, covid_pan) %>% summarise_mean_sd()) %>% 
  mutate(covid_pan = fct(covid_pan, c("pre_covid", "dur_covid", "total_dur")),
         icd_f50 = fct(icd_f50, c("F50-", "F50+", "all_diag"))) %>% 
  arrange(icd_f50, covid_pan)
df_ed_stats
```

## trend of ED diagnosis proportion
```{r}
df_ed %>% 
  arrange(adm_date) %>% 
  mutate(adm_month = as_factor(str_c(year(adm_date), month(adm_date), sep = ".")),
         adm_quarter = as_factor(quarter(adm_date, type = "year.quarter"))) %>%
  group_by(adm_month) %>% 
  summarise(#F50_prop = sum(icd_f50 == "F50+")/n()*100, 
            AN_prop = sum(f50_type == "Anorexie")/n()*100,
            BN_prop = sum(f50_type == "Bulimie")/n()*100,
            ED_sonst_prop = sum(f50_type == "Essstörung Sonst.")/n()*100) %>% 
  pivot_longer(ends_with("_prop"), names_to = "f50_prop_type", values_to = "f50_props") %>% 
  ggplot(aes(x = adm_month, y = f50_props, group = f50_prop_type)) + 
  geom_col(aes(fill = f50_prop_type)) +
  # geom_line(aes(colour = f50_prop_type)) +
  guides(x = guide_axis(angle = -90)) + 
  geom_vline(aes(xintercept = "2020.3"), show.legend = FALSE) +
  labs(title = "Proportion of cases with F50.* diagnoses per month", 
       x = "Year.Month", y = "cases with F50.* diagnosis [%]")
```

## periods of cases per patient
```{r}
df_ed %>% 
  group_by(p_id) %>%
  mutate(n_cases = n()) %>%
  filter(n_cases > 1) %>%
  # filter(case_id != first(case_id)) %>% 
  ungroup() %>%
  # slice_head(n = 500) %>%
  ggplot(aes(x = adm_date, xend = dis_date,
             y = fct_reorder(p_id, adm_date, first), yend = p_id,
             group = case_id, colour = f50_type)) +
  geom_segment() + 
  scale_x_datetime(date_breaks = "3 months", date_labels = "%y.%m", date_minor_breaks = "1 month", 
                   guide = guide_axis(angle = -90)) +
  # scale_y_discrete() +
  # scale_colour_viridis_d() + 
  geom_vline(aes(xintercept = as.POSIXct(covid_start)), show.legend = FALSE) + 
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) + 
  labs(title = "periods of treatments per patient (excl. first visit)", 
       x = "time", y = "patients", colour = "ED diag")
```

## counts & proportions
```{r}
gen_tabyl <- function(df, ...){
  df %>% 
    tabyl(..., show_missing_levels = FALSE) %>% 
    adorn_totals(c("row", "col")) %>% 
    adorn_percentages("row") %>%
    adorn_pct_formatting(digits = 1) %>%
    adorn_ns() %>% 
    adorn_title("combined")
}

df_ed %>% gen_tabyl(covid_pan, icd_f50)
df_ed %>% gen_tabyl(icd_f50, sex)
df_ed %>% gen_tabyl(icd_f50, sex, covid_pan)
```

## counts & proportions for ED subcategories
```{r}
df_ed %>% gen_tabyl(covid_pan, f50_type)
df_ed %>% gen_tabyl(f50_type, sex)
df_ed %>% gen_tabyl(f50_type, sex, covid_pan)
```
