---
title: "03_eating_disorder_stats"
author: "SP"
date: "`r Sys.Date()`"
output: 
  workflowr::wflow_html: 
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    df_print: paged
    fig_width: 11
    fig_height: 8
    fig_caption: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = FALSE)
# knitr::opts_knit$set(root.dir = dirname(getwd()))
```

```{r libs, include=FALSE}
do_source_data_creation <- FALSE
source("code/eating_disorder_data_proc.R")
load_inst_pkgs("tidyverse", "psych", "lmtest", "janitor", "knitr")
```

# Modelling
```{r echo=TRUE}
n_tests <- 39
```

## Essstörungen während Pandemie {.tabset .tabset-pills}
### Fälle mit ED-Diagnose
```{r echo=TRUE}
mod_ed_pan <- kruskal.test(icd_f50 ~ covid_pan, df_ed)
# mod_ed_pan <- tabyl(df_ed, icd_f50, covid_pan) %>% chisq.test()
mod_ed_pan
```
Corrected p-value for multiple testing is: `r format(mod_ed_pan[["p.value"]]*n_tests, scientific = FALSE)`

### ED-Subgruppe Anorexie
```{r echo=TRUE}
mod_an_pan <- kruskal.test(f50_type == "Anorexie" ~ covid_pan, df_ed)
mod_an_pan
```
Corrected p-value for multiple testing is: `r format(mod_an_pan[["p.value"]]*n_tests, scientific = FALSE)`

### ED-Subgruppe Bulimie
```{r echo=TRUE}
mod_bu_pan <- kruskal.test(f50_type == "Bulimie" ~ covid_pan, df_ed)
mod_bu_pan
```
Corrected p-value for multiple testing is: `r format(mod_bu_pan[["p.value"]]*n_tests, scientific = FALSE)`

### ED-Subgruppe sonstige Esstörungen
```{r echo=TRUE}
mod_edsonst_pan <- kruskal.test(f50_type == "Essstörung Sonst." ~ covid_pan, df_ed)
mod_edsonst_pan
```
Corrected p-value for multiple testing is: `r format(mod_edsonst_pan[["p.value"]]*n_tests, scientific = FALSE)`


## klinische Präsentation während Pandemie {.tabset .tabset-pills}
### jüngeres Alter
```{r echo=TRUE}
# kruskal.test(age_adm ~ covid_pan, df_ed)
mod_age_pan <- oneway.test(age_adm ~ covid_pan, df_ed)
mod_age_pan
```
Corrected p-value for multiple testing is: `r format(mod_age_pan[["p.value"]]*n_tests, scientific = FALSE)`

### Geschlecht
```{r echo=TRUE}
# kruskal.test(age_adm ~ covid_pan, df_ed)
mod_sex_pan <- kruskal.test(sex ~ covid_pan, df_ed)
mod_sex_pan
```
Corrected p-value for multiple testing is: `r format(mod_sex_pan[["p.value"]]*n_tests, scientific = FALSE)`

### Zunahme Komorbiditäten - F
```{r echo=TRUE}
mod_como_f_pan <- oneway.test(n_icd_f ~ covid_pan, df_ed)
mod_como_f_pan
```
Corrected p-value for multiple testing is: `r format(mod_como_f_pan[["p.value"]]*n_tests, scientific = FALSE)`

### Zunahme Komorbiditäten - !F
```{r echo=TRUE}
mod_como_other_pan <- oneway.test(n_icd_other ~ covid_pan, df_ed)
mod_como_other_pan
```
Corrected p-value for multiple testing is: `r format(mod_como_other_pan[["p.value"]]*n_tests, scientific = FALSE)`

### Abnahme (!) Behandlungsdauer
```{r echo=TRUE}
mod_los_pan <- oneway.test(length_stay_brutto ~ covid_pan, df_ed)
mod_los_pan
```
Corrected p-value for multiple testing is: `r format(mod_los_pan[["p.value"]]*n_tests, scientific = FALSE)`

### re-admissions (v1)
```{r}
# non-ED
chisq.test(matrix(c(226, 225, 575, 576), nrow = 2))
# ED
chisq.test(matrix(c(39, 63, 54, 98), nrow = 2))
# combined
chisq.test(matrix(c(226+39, 225+63, 575+54, 576+98), nrow = 2))
```

### re-adm 6 months
```{r}
kruskal.test(re_adm_soon_nona ~ covid_pan, df_ed)
kruskal.test(re_adm_soon_nona ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
kruskal.test(re_adm_soon_nona ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### age
```{r}
oneway.test(age_adm ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
oneway.test(age_adm ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### sex
```{r}
kruskal.test(sex ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
kruskal.test(sex ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### n diagnoses
```{r}
oneway.test(n_icd_code ~ covid_pan, df_ed)
oneway.test(n_icd_code ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
oneway.test(n_icd_code ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### somatic comorbidities
```{r}
oneway.test(n_icd_other ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
oneway.test(n_icd_other ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### psychiatric comorbidities
```{r}
oneway.test(n_icd_f ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
oneway.test(n_icd_f ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### LOS
```{r}
oneway.test(length_stay_brutto ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
oneway.test(length_stay_brutto ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
oneway.test(length_stay_brutto ~ covid_pan, df_ed %>% filter(icd_f50_dd_ad))
```

### DD
```{r}
kruskal.test(icd_dd ~ covid_pan, df_ed)
kruskal.test(icd_dd ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
kruskal.test(icd_dd ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### AD
```{r}
kruskal.test(icd_ad ~ covid_pan, df_ed)
kruskal.test(icd_ad ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
kruskal.test(icd_ad ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### F50+DD+AD
```{r}
kruskal.test(icd_f50_dd_ad ~ covid_pan, df_ed)
```

### psychotropic discharge medication
```{r}
kruskal.test(psych_med ~ covid_pan, df_ed)
kruskal.test(psych_med ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
kruskal.test(psych_med ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```


## Lockdowns

### Lockdown severity
```{r}
kruskal.test(icd_f50 ~ covid_lockd_sev, df_ed_covid)
# kruskal.test(f50_type ~ covid_lockd, df_ed_covid)
kruskal.test(f50_type == "Anorexie" ~ covid_lockd_sev, df_ed_covid)
kruskal.test(f50_type == "Bulimie" ~ covid_lockd_sev, df_ed_covid)
kruskal.test(f50_type == "Essstörung Sonst." ~ covid_lockd_sev, df_ed_covid)

kruskal.test(icd_f50_dd_ad ~ covid_lockd_sev, df_ed_covid)
```

### Covid periods
#### increasing trend of F50 share over periods
```{r}
kruskal.test(icd_f50 ~ covid_period_i, df_ed_covid)
cor.test(df_ed_covid$icd_f50 %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")$p.value

cor.test((df_ed_covid$f50_type == "Anorexie") %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")$p.value

cor.test((df_ed_covid$f50_type == "Bulimie") %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")$p.value

cor.test((df_ed_covid$f50_type == "Essstörung Sonst.") %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")$p.value

cor.test(df_ed_covid$icd_f50_dd_ad %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")$p.value
```

#### comparison with respective baseline period
```{r}
create_covid_period_baseline_df <- function(i = 1L) {
  df_ed %>% 
    # baseline period
    filter(covid_pan == "pre_covid", 
           as_date(adm_date_year0) %in% (df_covid_periods_is_lockd %>% 
                                           filter(period_i_is_lockd == i) %>% 
                                           use_series(period_days) %>% 
                                           extract2(1))) %>% 
    # covid_period
    bind_rows(df_ed %>% filter(as.numeric(covid_period_i) == i+1))
}

kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(1))
kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(2))
kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(3))
kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(4))
kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(5))
```


### Granger causality
```{r}
grangertest(AN_prop ~ as.numeric(covid_lockd_sev), order = 6, data = df_ed_covid_monthly)
# plot lag time points: orthogonal impulse response
```

```{r}
# granger plots
df_ed_covid_monthly %>% 
  transmute(adm_year_month, AN_prop, covid_lockd_sev = as.numeric(covid_lockd_sev)/15) %>% 
  pivot_longer(cols = -adm_year_month) %>% 
  ggplot(aes(x = adm_year_month, y = value, group = name)) +
  geom_line(aes(colour = name), linewidth = 0.7) +
  scale_colour_manual(values = c("#00AFBB", "#E7B800")) +
  theme_minimal()

library(tseries)
adf.test(df_ed_covid_monthly$AN_prop, k = 6)

library(vars)

ed_covid_monthly_irf <- df_ed_covid_monthly %>% 
  transmute(AN_prop, sev = as.numeric(covid_lockd_sev)) %>% 
  as.ts() %>% 
  VAR(p = 6, type = "const") %>% 
  irf()

plot(ed_covid_monthly_irf)
```


## GLM
! WIP, not finalised
```{r echo=TRUE}
# predicting ED
tmp <- glm(icd_f50 ~ covid_lockd_sev + sex + age_adm + n_icd_f + n_icd_other + length_stay_brutto,
           family = binomial,
           data = df_ed)
summary(tmp)
# sex
# age_adm

# f50_type
# length_stay_netto/brutto
# n_icd_f/other/code
```
